A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-03-27 reports/paper-rda.Rnw
Updated Rnw file for paper to summarize automatically the classification results in tables.

@@

In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}. We did not include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to avoid its excessive computations.  We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set. For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data. We then filtered the test data set with the corresponding $d$ features. After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier. Repeating this process 500 times, we computed the average of the error-rate estimates for each classifier. We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.  In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.

@@

 % latex table generated in R 2.15.1 by xtable 1.7-0 package
 % Sun Mar 24 19:54:37 2013
 \begin{table}[ht]
 \begin{center}
 \begin{tabular}{lrrrl}
   \hline
 Author & $N$ & $p$ & $K$ & Disease \\
   \hline
 Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
    \hline
 \end{tabular}
 \caption{Summary of high-dimensional microarray data sets.}
 \label{tab:data-summary}
 \end{center}
 \end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:20 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.232 (0.019) & 0.179 (0.017) & 0.513 (0.022) & 0.232 (0.019) & 0.231 (0.019) & 0.232 (0.019) \\ 
  Nakayama & 0.262 (0.020) & 0.308 (0.021) & 0.586 (0.022) & 0.262 (0.020) & 0.262 (0.020) & 0.278 (0.020) \\ 
  Shipp & 0.136 (0.015) & 0.106 (0.014) & 0.310 (0.021) & 0.136 (0.015) & 0.136 (0.015) & 0.136 (0.015) \\ 
  Singh & 0.111 (0.014) & 0.097 (0.013) & 0.338 (0.021) & 0.111 (0.014) & 0.111 (0.014) & 0.111 (0.014) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 100$. Approximate standard errors are given in parentheses.}
\label{tab:results-d100}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:37 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.244 (0.019) & 0.163 (0.017) & 0.459 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.243 (0.019) \\ 
  Nakayama & 0.244 (0.019) & 0.267 (0.020) & 0.498 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.259 (0.020) \\ 
  Shipp & 0.153 (0.016) & 0.104 (0.014) & 0.188 (0.017) & 0.153 (0.016) & 0.153 (0.016) & 0.154 (0.016) \\ 
  Singh & 0.135 (0.015) & 0.117 (0.014) & 0.188 (0.017) & 0.135 (0.015) & 0.135 (0.015) & 0.135 (0.015) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 250$. Approximate standard errors are given in parentheses.}
\label{tab:results-d250}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:48 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.243 (0.019) & 0.159 (0.016) & 0.429 (0.022) & 0.243 (0.019) & 0.243 (0.019) & 0.243 (0.019) \\ 
  Nakayama & 0.247 (0.019) & 0.248 (0.019) & 0.465 (0.022) & 0.247 (0.019) & 0.248 (0.019) & 0.261 (0.020) \\ 
  Shipp & 0.169 (0.017) & 0.097 (0.013) & 0.119 (0.015) & 0.169 (0.017) & 0.169 (0.017) & 0.170 (0.017) \\ 
  Singh & 0.174 (0.017) & 0.125 (0.015) & 0.155 (0.016) & 0.174 (0.017) & 0.174 (0.017) & 0.174 (0.017) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 500$. Approximate standard errors are given in parentheses.}
\label{tab:results-d500}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:05:01 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.249 (0.019) & 0.168 (0.017) & 0.381 (0.022) & 0.249 (0.019) & 0.249 (0.019) & 0.249 (0.019) \\ 
  Nakayama & 0.237 (0.019) & 0.238 (0.019) & 0.460 (0.022) & 0.237 (0.019) & 0.237 (0.019) & 0.246 (0.019) \\ 
  Shipp & 0.196 (0.018) & 0.106 (0.014) & 0.095 (0.013) & 0.196 (0.018) & 0.195 (0.018) & 0.198 (0.018) \\ 
  Singh & 0.224 (0.019) & 0.140 (0.016) & 0.142 (0.016) & 0.224 (0.019) & 0.224 (0.019) & 0.224 (0.019) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 1000$. Approximate standard errors are given in parentheses.}
\label{tab:results-d1000}
\end{center}
\end{table}

>>>>

<<setup, include=FALSE, cache=FALSE, echo=FALSE, warning=FALSE>>=
opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE, warning = FALSE,
               cache = FALSE, echo = FALSE)

setwd("..")
library(ProjectTemplate)
load.project()

# Clean up results for display
results <- na.omit(results)
results$d <- as.factor(results$d)
results$data_set <- as.character(results$data_set)

@

@@

In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}. We did not include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to avoid its excessive computations.  We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set. For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data. We then filtered the test data set with the corresponding $d$ features. After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier. Repeating this process 1000 times, we computed the average of the error-rate estimates for each classifier. We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.  In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.

@@

<<data_summary, results="asis">>=
# Table containing a summary of the data sets
data_summary <- subset(describe_data(), author %in% unique(results$data_set))
data_summary$K[data_summary$author == "nakayama"] <- 5
rownames(data_summary) <- NULL
colnames(data_summary)[1:3] <- c("Author", "Year", "N")
data_summary$Author <- Hmisc:::capitalize(as.character(data_summary$Author))
data_summary$Author <- with(data_summary, paste0(Author, " et al. (", Year, ")"))
data_summary <- subset(data_summary, select = -Year)
print(xtable(data_summary, caption = "Summary of high-dimensional microarray data sets.",
             label = "tab:data-summary", digits = 0), include.rownames = FALSE)

@

<<microarray_results, results="asis">>=
results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
                         Avg_Error = mean(Error),
                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))

# Tables containing the classification results for each value of 'q'
results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
results_summary$d <- as.factor(results_summary$d)
results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")

results_tables <- tapply(seq_along(results_summary$d), results_summary$d, function(i) {
  dcast(results_summary[i, ], Dataset ~ Classifier, value.var = "Error")
})

results_xtables <- lapply(names(results_tables), function(d) {
  tab_caption <- "The average of the test error rates for the microarray data sets with $d = "
  tab_caption <- paste0(tab_caption, d, "$. Approximate standard errors are given in parentheses.")
  xtable(results_tables[[d]], caption = tab_caption, label = paste0("tab:results-d", d))
})

lapply(results_xtables, print, include.rownames = FALSE)

@
